Manuscript

### 2018-12-20 manuscript_analysis.Rmd

g2p_dgls <-  unique(g2p_filt$Drug_Gene_Lin[!is.na(g2p_filt$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p_filt$Drug_Gene[!is.na(g2p_filt$Drug_Gene)])
g2p_genes <- unique(sapply(strsplit(g2p_druggenes, "_", fixed = TRUE), function(x) x[2]))
g2p_drugs <- unique(sapply(strsplit(g2p_druggenes, "_", fixed = TRUE), function(x) x[1]))
g2p_druggenes
g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

@@

@@

```{r}
MakeCorrResSummaryTable <- function(df_list, dataset_name) {
  res_full <- data.frame(matrix(ncol = 8, nrow = 0))
  for(i in 1:length(df_list)) {
    corr_res <- df_list[[i]] %>% group_by(G2P_DGL) %>% summarize(Dataset = dataset_name,
                Drug_Gene = names(df_list)[i],
                Lineages = paste(as.character(unique(group_general_lineage_name)), collapse = ", "),
                N_Lineages = length(as.character(unique(group_general_lineage_name))),
                Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$estimate,
                Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs")$estimate)
    corr_res <- merge(corr_res, compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, data = df_list[[i]], group.by = "G2P_DGL", method = "wilcox.test")[, c("G2P_DGL", "p")], by = "G2P_DGL", all = TRUE)
    corr_res$Mut_Signif <- ifelse(corr_res$p < 0.05, 1, 0)
    res_full <- rbind(res_full, corr_res)
  }
  return(res_full)
}
```

@@

g2p_corr_all <- rbind(ccle_corr_all, ctrp_corr_all, gdsc_corr_all)
```

```{r}
G2PCorr_KStest <- function(dataset, dataset_name) {
  Metric <- colnames(dataset)[6:9]
  D_stat <- NULL
  pval <- NULL
  signif <- NULL

```{r}
ccle_corr_stats <- G2PCorr_KStest(ccle_corr_all, "CCLE")
ctrp_corr_stats <- G2PCorr_KStest(ctrp_corr_all, "CTRP")
gdsc_corr_stats <- G2PCorr_KStest(gdsc_corr_all, "GDSC")
g2p_corr_stats <- rbind(ccle_corr_stats, ctrp_corr_stats, gdsc_corr_stats)
g2p_corr_stats

@@

@@

>>>>

g2p_dgls <-  unique(g2p_filt$Drug_Gene_Lin[!is.na(g2p_filt$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p_filt$Drug_Gene[!is.na(g2p_filt$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
g2p_druggenes
g2p_drugs
g2p_genes
g2p_lins

# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

@@

Pearson correlations for gene expression and copy number:
```{r fig.height=6, fig.width=12}
crispr_pearson_res_GE <- transform(crispr_pearson_res_GE, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_pearson_res_GE$p_log10 <- -log10(crispr_pearson_res_GE$p)
crispr_pearson_res_CN <- transform(crispr_pearson_res_CN, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_pearson_res_CN$p_log10 <- -log10(crispr_pearson_res_CN$p)

crispr_ge_pearson_corr_plot <- ggplot(data = crispr_pearson_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.65, 0.25)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Pearson Correlation Coefficient (r_s)", subtitle = "CRISPR: Correlation between AUC z-score and gene expression")
crispr_cn_pearson_corr_plot <- ggplot(data = crispr_pearson_res_CN, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.65, 0.25)) +
  theme(legend.position = "bottom", axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Pearson Correlation Coefficient (r_s)", subtitle = "CRISPR: Correlation between AUC z-score and copy number")

crispr_pearson_corr_plot <- plot_grid(crispr_ge_pearson_corr_plot, crispr_cn_pearson_corr_plot, align = "h")
crispr_pearson_corr_plot
# ggsave("./plots/manuscript/crispr_pearson_corr_plot.png", crispr_pearson_corr_plot, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

Spearman correlations for gene expression and copy number:
```{r fig.height=6, fig.width=12}
crispr_spearman_res_GE <- transform(crispr_spearman_res_GE, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_GE$p_log10 <- -log10(crispr_spearman_res_GE$p)
crispr_spearman_res_CN <- transform(crispr_spearman_res_CN, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_CN$p_log10 <- -log10(crispr_spearman_res_CN$p)

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)", subtitle = "CRISPR: Correlation between AUC z-score and gene expression")
crispr_cn_spearman_corr_plot <- ggplot(data = crispr_spearman_res_CN, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)", subtitle = "CRISPR: Correlation between AUC z-score and copy number")

crispr_spearman_corr_plot <- plot_grid(crispr_ge_spearman_corr_plot, crispr_cn_spearman_corr_plot, align = "h")
crispr_spearman_corr_plot
# ggsave("./plots/manuscript/crispr_spearman_corr_plot.png", crispr_spearman_corr_plot, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

@@

```{r}
MakeCorrResSummaryTable <- function(df_list, dataset_name) {
  res_full <- data.frame(matrix(ncol = 8, nrow = 0))
  for(i in 1:length(df_list)) {
    corr_res <- df_list[[i]] %>% group_by(G2P_DGL) %>% summarize(Dataset = dataset_name,
                Drug_Gene = names(df_list)[i],
                Lineages = paste(as.character(unique(group_general_lineage_name)), collapse = ", "),
                N_Lineages = length(as.character(unique(group_general_lineage_name))),
                Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$estimate,
                Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs")$estimate)
    corr_res <- merge(corr_res, compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, data = df_list[[i]], group.by = "G2P_DGL", method = "wilcox.test")[, c("G2P_DGL", "p")], by = "G2P_DGL", all = TRUE)
    corr_res$Mut_Signif <- ifelse(corr_res$p < 0.05, 1, 0)
    res_full <- rbind(res_full, corr_res)
  }
  return(res_full)
}

G2PCorr_KStest <- function(dataset, dataset_name) {
  Metric <- colnames(dataset)[6:9]
  D_stat <- NULL
  pval <- NULL
  signif <- NULL

@@

g2p_corr_all <- rbind(ccle_corr_all, ctrp_corr_all, gdsc_corr_all)

ccle_corr_stats <- G2PCorr_KStest(ccle_corr_all, "CCLE")
ctrp_corr_stats <- G2PCorr_KStest(ctrp_corr_all, "CTRP")
gdsc_corr_stats <- G2PCorr_KStest(gdsc_corr_all, "GDSC")
g2p_corr_stats <- rbind(ccle_corr_stats, ctrp_corr_stats, gdsc_corr_stats)
g2p_corr_stats <- g2p_corr_stats %>% separate(col = Metric, into = c("Correlation_Method", "corr", "Metric"), sep = "_")
g2p_corr_stats

@@

Plot of Kolmogorov-Smirnov test results comparing correlations between CN/GE and AUC z-score in G2P and non-G2P lineages:
```{r fig.height=6, fig.width=12}
g2p_corr_4plot <- g2p_corr_all
g2p_corr_4plot$p <- NULL
g2p_corr_4plot$Mut_Signif <- NULL
g2p_corr_4plot <- melt(g2p_corr_4plot, id.vars = colnames(g2p_corr_4plot)[1:5], measure.vars = colnames(g2p_corr_4plot)[6:9], value.name = "Correlation")
g2p_corr_4plot <- g2p_corr_4plot %>% separate(col = variable, into = c("Correlation_Method", "corr", "Metric"), sep = "_", remove = FALSE)

g2p_ks_res_plot <- ggplot(data = g2p_corr_4plot) +
  facet_grid(Correlation_Method ~ Metric * Dataset) +
  geom_violin(aes(y = Correlation, x = G2P_DGL, fill = G2P_DGL, color = G2P_DGL), alpha = 0.3) +
  geom_boxplot(width = 0.075, aes(y = Correlation, x = G2P_DGL, color = G2P_DGL)) +
  geom_jitter(width = 0.3, alpha = 0.7, aes(y = Correlation, x = G2P_DGL, color = G2P_DGL)) +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(-1, 1)) +
  geom_text(data = g2p_corr_stats, aes(x = 1.5, y = 0.8, label = paste0("D = ", formatC(D_stat, digits = 2, format = "f"), ", p = ", ifelse(pval > 0.1, round(pval, 2), round(pval, 3)), "\n", ifelse(is.na(signif), "", signif)))) +
  labs(x = "Is the lineage in a G2P DGL?", y = "Correlation Coefficint", subtitle = "Kolmogorov-Smirnov test results comparing correlations between CN/GE and AUC z-score in G2P and non-G2P lineages.")
g2p_ks_res_plot

# ggsave("./plots/manuscript/g2p_ks_res_plot.png", g2p_ks_res_plot, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

@@

## Pancancer vs. Lineage

```{r}
compare_grid <- expand.grid(g2p_druggenes, unique(ccl_converter$group_general_lineage_name), c("CCLE", "CTRP", "GDSC"), c("Gene Expression", "Copy Number", "Mutation Status"))
colnames(compare_grid) <- c("Drug_Gene", "group_general_lineage_name", "Dataset", "Metric")

pancan_ge <- select(filter(all_pearson_res, Metric == "Gene Expression"), Drug_Gene, p, Dataset, p_log10)
colnames(pancan_ge)[colnames(pancan_ge) == "p"] <- "p_pancan"
colnames(pancan_ge)[colnames(pancan_ge) == "p_log10"] <- "p_log10_pancan"
compare_ge <- select(dgl_signif_pearson_g2p_grid, Drug_Gene, group_general_lineage_name, n, Pearson_Pval_GE, Drug_Gene_Lin, InG2P, Dataset)
colnames(compare_ge)[colnames(compare_ge) == "Pearson_Pval_GE"] <- "p"
compare_ge$p_log10 <- -log10(compare_ge$p)
compare_ge <- merge(compare_ge, pancan_ge, by = c("Drug_Gene", "Dataset"), all = TRUE)
compare_ge$Compare_Pvals <- ifelse(compare_ge$p_log10 < compare_ge$p_log10_pancan, "Under", "Above")
compare_ge$Metric <- "Gene Expression"

pancan_cn <- select(filter(all_pearson_res, Metric == "Copy Number"), Drug_Gene, p, Dataset, p_log10)
colnames(pancan_cn)[colnames(pancan_cn) == "p"] <- "p_pancan"
colnames(pancan_cn)[colnames(pancan_cn) == "p_log10"] <- "p_log10_pancan"
compare_cn <- select(dgl_signif_pearson_g2p_grid, Drug_Gene, group_general_lineage_name, n, Pearson_Pval_CN, Drug_Gene_Lin, InG2P, Dataset)
colnames(compare_cn)[colnames(compare_cn) == "Pearson_Pval_CN"] <- "p"
compare_cn$p_log10 <- -log10(compare_cn$p)
compare_cn <- merge(compare_cn, pancan_cn, by = c("Drug_Gene", "Dataset"), all = TRUE)
compare_cn$Compare_Pvals <- ifelse(compare_cn$p_log10 < compare_cn$p_log10_pancan, "Under", "Above")
compare_cn$Metric <- "Copy Number"

pancan_mut <- all_mut_res
pancan_mut$Metric <- NULL
pancan_mut$corr <- NULL
colnames(pancan_mut)[colnames(pancan_mut) == "p"] <- "p_pancan"
colnames(pancan_mut)[colnames(pancan_mut) == "p_log10"] <- "p_log10_pancan"
compare_mut <- dgl_mut_grid_summ
compare_mut$p_log10 <- -log10(compare_mut$p)
compare_mut <- merge(compare_mut, pancan_mut, by = c("Drug_Gene", "Dataset"), all = TRUE)
compare_mut$Compare_Pvals <- ifelse(compare_mut$p_log10 < compare_mut$p_log10_pancan, "Under", "Above")
compare_mut$Metric <- "Mutation Status"

compare_data <- filter(rbind(compare_ge, compare_cn, compare_mut), Drug_Gene %in% g2p_druggenes)
compare_data <- merge(compare_grid, compare_data, by = c("Drug_Gene", "group_general_lineage_name", "Dataset", "Metric"), all.x = TRUE)
```

```{r}
compare_data_stats <- compare_data %>% count(Metric, Compare_Pvals)
compare_data_stats
```

```{r fig.height=12, fig.width=12}
compare_mut_plot <- ggplot(data = filter(compare_data, Metric == "Mutation Status"), aes(x = group_general_lineage_name, y = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(aes(shape = Compare_Pvals, fill = p_log10), color = "gray50") +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  scale_shape_manual(values = c(25, 24), labels = c("Below", "Above")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4), legend.position = "top") +
  labs(title = "Mutation Status",
       subtitle = paste0("Above: ", filter(compare_data_stats, Metric == "Mutation Status" & Compare_Pvals == "Above")$nn, "\n",
                         "Under: ", filter(compare_data_stats, Metric == "Mutation Status" & Compare_Pvals == "Under")$nn, "\n",
                         "Not tested: ", filter(compare_data_stats, Metric == "Mutation Status" & is.na(Compare_Pvals))$nn),
       x = "Lineage", y = "Drug-Gene", shape = "Compare linege p-value to pancancer p-value:", fill = "-log10(p)")
# ggsave("./plots/manuscript/compare_mut_plot.png", compare_mut_plot, device = "png", dpi = 450, width = 12, height = 12, units = "in")

compare_ge_plot <- ggplot(data = filter(compare_data, Metric == "Gene Expression"), aes(x = group_general_lineage_name, y = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(aes(shape = Compare_Pvals, fill = p_log10), color = "gray50") +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  scale_shape_manual(values = c(25, 24), labels = c("Below", "Above")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4), legend.position = "top") +
  labs(title = "Gene Expression",
       subtitle = paste0("Above: ", filter(compare_data_stats, Metric == "Gene Expression" & Compare_Pvals == "Above")$nn, "\n",
                         "Under: ", filter(compare_data_stats, Metric == "Gene Expression" & Compare_Pvals == "Under")$nn, "\n",
                         "Not tested: ", filter(compare_data_stats, Metric == "Gene Expression" & is.na(Compare_Pvals))$nn),
       x = "Lineage", y = "Drug-Gene", shape = "Compare linege p-value to pancancer p-value:", fill = "-log10(p)")
# ggsave("./plots/manuscript/compare_ge_plot.png", compare_ge_plot, device = "png", dpi = 450, width = 12, height = 12, units = "in")

compare_cn_plot <- ggplot(data = filter(compare_data, Metric == "Copy Number"), aes(x = group_general_lineage_name, y = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(aes(shape = Compare_Pvals, fill = p_log10), color = "gray50") +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  scale_shape_manual(values = c(25, 24), labels = c("Below", "Above")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4), legend.position = "top") +
  labs(title = "Copy Number",
       subtitle = paste0("Above: ", filter(compare_data_stats, Metric == "Copy Number" & Compare_Pvals == "Above")$nn, "\n",
                         "Under: ", filter(compare_data_stats, Metric == "Copy Number" & Compare_Pvals == "Under")$nn, "\n",
                         "Not tested: ", filter(compare_data_stats, Metric == "Copy Number" & is.na(Compare_Pvals))$nn),
       x = "Lineage", y = "Drug-Gene", shape = "Compare linege p-value to pancancer p-value:", fill = "-log10(p)")
# ggsave("./plots/manuscript/compare_cn_plot.png", compare_cn_plot, device = "png", dpi = 450, width = 12, height = 12, units = "in")
```

```{r}
test <- filter(compare_mut, Drug_Gene == "nilotinib_ABL1" & Dataset == "CCLE" & group_general_lineage_name == "bone cancer")
```
